\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-06-22 coqimpl.tex
minimal edits in coqimpl.tex

After considering the pros and cons, we decide to implement a Coq Plugin in which we can add new Vernacular command and translate each new command into a bunch of Coq commands (the surface Vernacular command) on the fly, instead of modifying the code-base of Coq.

Despite of the possible difficulty of the future maintenance, this approach has various advantages: 1. it is the easiest and the most accessible way to prototype as it relieves us the necessity of familarity of Coq base, especially for the implementation of the module and functor; 2. we have a clear definition of trusted-base---the whole Coq; 3. it is easy to debug---we just need to check the translated commands; 4. it can be well-incorporated with the existent tools like VSCoq; 5. it is more accessible for the the interested audience who can then easily adapt our plugin and give a try---otherwise they have to download and re-build the whole customized Coq; 6. it is more stable because the surface syntax of Coq should be stable across different versions; 7. still, this plugin can capture the key ingredients of implementing family polymorphism inside Coq and act as a reference for guiding an appropriate implementation of Family Polymorphism inside all sorts of proof assistants.

\textbf{Family compiled into module.} This is the main idea of compiling family components into Coq's primitives: we compile family into coq module, family type into module type and context of a given term into parameters of the module. Unlike module, we need to faithfully reflect the \textit{late binding} nature of family : for each defining field $ .. \cL\sigma \vdash t : T $, we need to compile $t$ using ``universal-quantifier-wrapped term'' : $.. \vdash \lambda t : \forall \cL \sigma. T$. However, instead of using Coq's universal quantifier, we choose module and module parameters to achieve this wrapping---each family type in the context of the judgement will be one module parameter with the compiled module type, and thus each field of the family will compile into a functor parametrized by its context. Doing so we can also get rapid feedback from type checking when the users are defining each field interactively with Coq.
% Insert one pseudo-code example for explanation of the mechanism
\begin{figure}
  \begin{minipage}[t]{0.25\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Family A.
Field a : nat 
  := 1.
Field b 
  : self_A.a = self_A.a
  := eq_refl.
EndFamily.
\end{minted}
  \end{minipage}
  \begin{minipage}[t]{0.4\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Module a_4 (self_A: EmptySig).
Definition a : nat := 1. End a_4.
Module Type a_5 (self_A: EmptySig).
Parameter (a : nat). End a_5.
Module Type a_6 := a_5 EmptyMod.
\end{minted}
  \end{minipage}
  \begin{minipage}[t]{0.3\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Module b_7 (self_A: a_6).
Definition b 
  : self_A.a = self_A.a 
  := eq_refl. End b_7.

@@

% Concrete details of compilation of a family term data structure
Taking \cref{fig:plugin-example1} as an example: when the user interactively input the line "Field a : nat := ..", our plugin will translate this statement into the "Module a_9", "Module Type a_10, a_11" three components. {The "EmptySig" is just empty module type.} Then, these modules and module types will be generated and type-checked by Coq immediately. The computed "Module a_10" is then part of the context of "Field b" (as the module type of "self__A") for "b" to refer to. The internal representation of family "A" is simply the list of these ``compiled'' modules and module types. We can simply achieve incremental checking and overriding---for example, we want to override "Field a", we simply replace that compiled module with the new module inside the list. This compiled module type serve as an abstraction to let overriding possible.

>>>>

After considering the pros and cons, we decide to implement a Coq Plugin in which we can add new Vernacular commands and translate each new command into a bunch of Coq commands (the surface Vernacular command) on the fly, instead of modifying the code-base of Coq.

Despite possible difficulties for future maintenance, this approach has
various advantages: 1. it is the easiest and the most accessible way to
prototype as it relieves us the necessity of familarity of Coq base,
especially for the implementation of the module and functor; 2. we have
a clear definition of trusted-base---the whole Coq; 3. it is easy to
debug---we just need to check the translated commands; 4. it can be
well-incorporated with the existent tools like VSCoq; 5. it is more
accessible for the interested audience who can then easily adapt our
plugin and give a try---otherwise they have to download and re-build the
whole customized Coq; 6. it is more stable because the surface syntax of
Coq should be stable across different versions; 7. still, this plugin
can capture the key ingredients of implementing family polymorphism
inside Coq and act as a reference for guiding an appropriate
implementation of Family Polymorphism inside all sorts of proof
assistants.\YZ{Does the Coq implementation perform type checking to
prevent illegal uses fam poly?}

\textbf{Family compiled into module.} This is the main idea of compiling
family components into Coq's primitives: we compile family into Coq
module, family type into module type, and the context of a given term
into parameters of the module. Unlike module, we need to faithfully
reflect the \textit{late binding} nature of family: for each defining
field $ .. \cL\sigma \vdash t : T $, we need to compile $t$ using
``universal-quantifier-wrapped term'' : $.. \vdash \lambda t : \forall
\cL \sigma. T$. However, instead of using Coq's universal quantifier, we
choose module and module parameters to achieve this wrapping---each
family type in the context of the judgement will be one module parameter
with the compiled module type, and thus each field of the family will
compile into a functor parametrized by its context. Doing so we can also
get rapid feedback from type checking when the users are defining each
field interactively with Coq.
% Insert one pseudo-code example for explanation of the mechanism
\begin{figure}[!htb]
  \begin{minipage}[t]{0.25\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Family A.
Field a : nat 
  := 1.
Field b 
  : self_A.a = self_A.a
  := eq_refl.
EndFamily.
\end{minted}
  \end{minipage}\YZ{Can the self's be omitted?}
  \begin{minipage}[t]{0.4\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Module a_4 (self_A: EmptySig).
Definition a : nat := 1. End a_4.
Module Type a_5 (self_A: EmptySig).
Parameter (a : nat). End a_5.
Module Type a_6 := a_5 EmptyMod.
\end{minted}
  \end{minipage}
  \begin{minipage}[t]{0.3\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Module b_7 (self_A: a_6).
Definition b 
  : self_A.a = self_A.a 
  := eq_refl. End b_7.

@@

% Concrete details of compilation of a family term data structure
Taking \cref{fig:plugin-example1} as an example: when the user interactively inputs the line "Field a : nat := ..", our plugin will translate this statement into the "Module a_9", "Module Type a_10, a_11" three components. {The "EmptySig" is just empty module type.} Then, these modules and module types will be generated and type-checked by Coq immediately. The computed "Module a_10" is then part of the context of "Field b" (as the module type of "self__A") for "b" to refer to. The internal representation of family "A" is simply the list of these ``compiled'' modules and module types. We can simply achieve incremental checking and overriding---for example, we want to override "Field a", we simply replace that compiled module with the new module inside the list. This compiled module type serve as an abstraction to let overriding possible.
